### 2021-07-25 paper/main.tex
Initial commit

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[
% twocolumn,
% hf,
]{ceurart}

%%
%% One can fix some overfulls
% \sloppy

%%
%% Minted listings support 
%% Need pygment <http://pygments.org/> <http://pypi.python.org/pypi/Pygments>
\usepackage{minted}
%% auto break lines
\setminted{breaklines=true}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% Rights management information.
%% CC-BY is default license.
\copyrightyear{2021}
\copyrightclause{Copyright for this paper by its authors.
  Use permitted under Creative Commons License Attribution 4.0
  International (CC BY 4.0).}

%%
%% This command is for the conference information
\conference{Woodstock'21: Symposium on the irreproducible science,
  June 07--11, 2021, Woodstock, NY}

%%
%% The "title" command
\title{A better way to format your document for CEUR-WS}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
\author[1]{Robert Zinkov}[%
%orcid=0000-0002-0877-7063,
email=zinkov@robots.ox.ac.uk,
url=https://zinkov.com/,
]
\address[1]{Peoples' Friendship University of Russia (RUDN University),
  6 Miklukho-Maklaya St, Moscow, 117198, Russian Federation}

\author[2]{William E. Byrd}[%
%orcid=0000-0001-7116-9338,
email=webyrd@uab.edu,
url=http://webyrd.net/,
]
\address[2]{Vrije Universiteit Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, The Netherlands}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Probabilistic programming is sometimes described as a generalisation
  of logic programming where instead of just returning a set of facts
  that answer a given query, we return a probability distribution over
  facts that answer a given query. But many contemporary probabilistic
  logic programming languages are not implemented as simple extensions
  of existing logic programming languages but instead involve their
  own unique implementations. Here we introduce probKanren, a simple
  extension to miniKanren that transforms
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\begin{keywords}
  Probabilistic Logic Programming \sep
  miniKanren \sep
  Probabilistic Programming \sep
  Sequential Monte Carlo
  %Particle Filtering
  %relational programming
  %logic programming
\end{keywords}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

Conceptually, logic programming provides a way to model
non-determinism. This is accomplished by maintaining a set of answers
that satisfy a set of logical constraints. A natural generalisation
to this domain is adding a notion of uncertainty to this set of answers
by associating with them a probability distribution.

\subsection{Illustrated Example}

To help explain how to use probKanren we introduce the following example:

\begin{minted}{Scheme}
  (run 1000 (q)
    (normal 0 3 q)
    (greatero q 0))
\end{minted}

This probKanren program draws 1000 samples from a normal distribution truncated
below 0.

\subsection{Contributions}

\section{Related Work}

There is a rich history of extending logic programming formalisms to
support probabilistic inference. Early systems like
PRISM\cite{sato1997prism} and ProbLog\cite{de2007problog} allowed
associating discrete distributions with facts. Later
work\cite{gutmann2010extending} introduces distribution clauses so
that some continuous distributions. Other work extended these methods
further while focusing on efficient exact inference algorithms like
model weight integration\cite{islam2012inference,
  belle2015probabilistic}. Our work is most similar to
\cite{gutmann2011magic} except that while they combine their forward
reasoning with an importance sampler we use a particle cascade instead
which can be more sample efficient.

\section{Background}

\subsection{miniKanren}

\textsc{miniKanren}\cite{byrd2017unified, daniel2018reasoned} is a
pure logic programming language embedded in Scheme. The language
consists of a set of terms, functions for constructing, a set of goal
primitives, and two run functions to answer queries in the
language. The goal primitives consist of a \texttt{fresh} form for
introducing logic variables, a unification primitive, a conjunction
combinator, and a disjunction combinator.

\subsection{Probabilistic Programming}

Probabilistic Programming Languages\cite{wood2014new,
  van2018introduction} are a family of domain specific languages for
posing and efficiently solving probabilistic modelling problems. At
their core, all have a way to \texttt{sample} from a probability
distribution and \texttt{observe} data generated from a probability
distribution.

There are many ways to implement inference algorithms for probabilistic
programming languages but methods based on likelihood-weighting and
sequential monte carlo algorithms are the easiest.

\subsection{Sequential Monte Carlo}

Sequential Monte Carlo\cite{chopin2020introduction} are an efficient
online way to sample from probabilistic models especially suited for
state-space domains. If we imagine our probabilistic programs as
straight-line programs with no control-flow we can imagine numbering
every sample function $x_1, f_2, \ldots, f_n$ and every observe
function $g_1, g_2, \ldots, g_n$ then our probability density over our
random variables $x$ and observed data $y$ can be defined as:

\begin{equation}
  p(x, y) = \prod_i f(x_i \mid x_{0:n-1})g(y_i \mid x_{0:n})
\end{equation}

\subsection{Sequential Importance Sampling}

If we imagine running this program each execution trace can be seen as
a sample from the distribution. If we then weight by their likelihood
as we run them that collection execution traces will be an empirical
distribution of the given program. We call each of these execution traces
particles and the below method is how we obtain them along with their weights $w$.

\begin{equation}
  x_n \sim f(x_n \mid x_{0:n-1}) \\
  w_n^{(k)} = \frac{g(x_n \mid x_{0:n})f(x_n \mid x_{0:n-1})}{q(x_n \mid x_{0:n-1})} \\
  W_n^{(k)} = W_{n-1}^{(k)} w_n^{(k)}
\end{equation}

The weights are then normalised at the end to make a proper probability distribution.

\subsection{Sequential Importance Resampling}

The problem with the above algorithm is over time for many particles
$W_k$ is going to become low and that particle will stop being very
informative of the underlying distribution. To mitigate this issue,
each time we encounter an observation we resample our particles. Resampling
effectively removes particles with low weight and duplicates particles
with higher weight by sampling with replacement our existing particles.

The above is called multinomial resampling but there are other methods
as well. A survey\cite{douc2005comparison} of resampling methods
suggests all of them are helpful to reduce particle degeneracy.

\subsection{Particle Cascade}

As the SMC resampling step was defined in the previous section
Particle Cascades \cite{PaigeWDT14} remove this barrier allowing
every particle to be resampled asynchronously with the associated
weights being relative to a global running average. % double-check this

\section{Proposed Method}

We propose to extend \textsc{miniKanren} by augmenting each of the
search streams with a set of particles. These particles represent the
empirical distribution of that stream. Each particle has associated
with it a substitution of all the logic and random variables as well
as a weight that is proportional to the likelihood of the substitution.

An initial set of particles is created from the probabilistic program
when it is first run. As disjunctions like \texttt{conde} are encountered,
we split evenly the number of particles allocated to each stream. Whenever
we encounter a unification primitive, we run a resampling step. This
helps to prune low-weight particles and replicate high weight ones.

As an optimisation we may create more particles during resampling
based on a globally stored a counter of the effective sample size of
all particles across all streams.

We follow \cite{gutmann2010extending} and place the following restrictions
on our distribution clauses and the random variables they specify.

Firstly, the arguments of distribution clauses must be
ground. Secondly, a random variable can not unify with any arithmetic
expression

\section{Experiments}

One of the experiments should be something from the probabilistic logic literature.

Friends who Smoke is a probabilistic logic program which models the social nature of
who smokes cigarettes. The model predicts that people who are friends with people
who smoke are more likely to smoke.

We could include experiment from Noah Goodman's dippl work like a semantic parsing
example\\ \url{http://dippl.org/examples/semanticparsing.html}

We could include program synthesis experiment, where probabilities
allow us to specify a soft preference for using certain language
primitives in a similar spirit to
\textsc{rKanren}\cite{swords2013rkanren} or neural-guided search
\cite{zhang2018neural}.

\section{Conclusions}

We made a cool and simple to implement extension to
\textsc{miniKanren} that let's us support probabilistic inference on
both discrete and continuous distributions.

%% \begin{table*}
%%   \caption{Frequency of Special Characters}
%%   \label{tab:freq}
%%   \begin{tabular}{ccl}
%%     \toprule
%%     Non-English or Math&Frequency&Comments\\
%%     \midrule
%%     \O & 1 in 1,000& For Swedish names\\
%%     $\pi$ & 1 in 5& Common in math\\
%%     \$ & 4 in 5 & Used in business\\
%%     $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
%%   \bottomrule
%% \end{tabular}
%% \end{table*}

%% To set a wider table, which takes up the whole width of the page's
%% live area, use the environment \verb|table*| to enclose the table's
%% contents and the table caption.  As with a single-column table, this
%% wide table will ``float'' to a location deemed more
%% desirable. Immediately following this sentence is the point at which
%% Table~\ref{tab:commands} is included in the input file; again, it is
%% instructive to compare the placement of the table here with the table
%% in the printed output of this document.

%% \begin{table}
%%   \caption{Some Typical Commands}
%%   \label{tab:commands}
%%   \begin{tabular}{ccl}
%%     \toprule
%%     Command &A Number & Comments\\
%%     \midrule
%%     \texttt{{\char'134}author} & 100& Author \\
%%     \texttt{{\char'134}table}& 300 & For tables\\
%%     \texttt{{\char'134}table*}& 400& For wider tables\\
%%     \bottomrule
%%   \end{tabular}
%% \end{table}

%% The ``\verb|figure|'' environment should be used for figures. One or
%% more images can be placed within a figure. If your figure contains
%% third-party material, you must clearly identify it as such, as shown
%% in the example below.
%% \begin{figure}
%%   \centering
%%   %\includegraphics[width=\linewidth]{sample-franklin}
%%   \caption{1907 Franklin Model D roadster. Photograph by Harris \&
%%     Ewing, Inc. [Public domain], via Wikimedia
%%     Commons. (\url{https://goo.gl/VLCRBB}).}
%% \end{figure}

%% Define the bibliography file to be used
\bibliography{main}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\end{document}

### 2021-07-25 paper/main.tex
Add title
added Will's affiliation

@@
 %% The "title" command
-\title{A better way to format your document for CEUR-WS}
+\title{probKanren: A Simple Probabilistic extension for miniKanren}
 
@@
 ]
-\address[2]{Vrije Universiteit Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, The Netherlands}
+\address[2]{Hugh Kaul Precision Medicine Institute, University of Alabama at Birmingham, 705 20th Street S., Birmingham, AL 35233, United States of America}
 

### 2021-07-25 paper/main.tex
Add more TODO notes

@@
   of logic programming where instead of just returning a set of facts
-  that answer a given query, we return a probability distribution over
+  that answer a given query, which return a probability distribution over
   facts that answer a given query. But many contemporary probabilistic
@@
 
+Define terms, goals, and substitutions here
+
 \subsection{Probabilistic Programming}
@@
 
-\begin{equation}
-  x_n \sim f(x_n \mid x_{0:n-1}) \\
-  w_n^{(k)} = \frac{g(x_n \mid x_{0:n})f(x_n \mid x_{0:n-1})}{q(x_n \mid x_{0:n-1})} \\
-  W_n^{(k)} = W_{n-1}^{(k)} w_n^{(k)}
-\end{equation}
+\begin{eqnarray}
+  x_n &\sim& f(x_n \mid x_{0:n-1}) \\
+  w_n^{(k)} &=& \frac{g(x_n \mid x_{0:n})f(x_n \mid x_{0:n-1})}{q(x_n \mid x_{0:n-1})} \\
+  W_n^{(k)} &=& W_{n-1}^{(k)} w_n^{(k)}
+\end{eqnarray}
 
@@
 weights being relative to a global running average. % double-check this
+
+\section{Grammar and Definitions}
+
+Define distributional clauses here
 

### 2021-08-23 paper/main.tex
Add updated code to repo

@@
 \usepackage{minted}
+\usepackage{syntax}
+\shortverb{\|}
 %% auto break lines
@@
 %% This command is for the conference information
-\conference{Woodstock'21: Symposium on the irreproducible science,
-  June 07--11, 2021, Woodstock, NY}
+\conference{PLP '21: The 8th Workshop on Probabilistic Logic Programming,
+  September 20--27, 2021, Porto, Portugal}
 
@@
 %% The "title" command
-\title{probKanren: A Simple Probabilistic extension for miniKanren}
+\title{probKanren: A Simple Probabilistic extension for microKanren}
 
@@
 ]
-\address[1]{Peoples' Friendship University of Russia (RUDN University),
-  6 Miklukho-Maklaya St, Moscow, 117198, Russian Federation}
+\address[1]{Dept. of Engineering Science, University of Oxford,
+  25 Banbury Rd, Oxford, UK}
 
@@
   Probabilistic programming is sometimes described as a generalisation
-  of logic programming where instead of just returning a set of facts
-  that answer a given query, which return a probability distribution over
-  facts that answer a given query. But many contemporary probabilistic
-  logic programming languages are not implemented as simple extensions
-  of existing logic programming languages but instead involve their
-  own unique implementations. Here we introduce probKanren, a simple
-  extension to miniKanren that transforms
+  of logic programming where instead of just returning a set of
+  answers to a given query, we also return a probability distribution
+  over those answers. But many contemporary probabilistic logic
+  programming languages are not implemented as simple extensions of
+  existing logic programming languages but instead involve their own
+  unique implementations. Here we introduce probKanren, a simple
+  extension to microKanren that transforms it into a probabilistic
+  programming language without needing to make any modifications to
+  the underlying search.
 \end{abstract}
@@
 
+But the conceptual simplicity of this sort of generalisation is not reflected
+in the complexity of many existing probabilistic logic programming systems.
+They often involve making implementing sophisticated algorithms and the underlying
+systems are not just implemented on top of existing logic programming systems.
+
+We believe a conceptually simple extension deserves a conceptually
+simple implementation to go along with it. We thus contribute a simple
+way to extend microkanren a small logic programming DSL such that it
+becomes a probabilistic programming language.
+
 \subsection{Illustrated Example}
@@
 
+\begin{minted}{Scheme}
+  (run 1000 (q)
+    (beta 1 1 q)
+    (flip q #t))
+\end{minted}
+
 This probKanren program draws 1000 samples from a normal distribution truncated
 below 0.
-
-\subsection{Contributions}
 
@@
 PRISM\cite{sato1997prism} and ProbLog\cite{de2007problog} allowed
-associating discrete distributions with facts. Later
-work\cite{gutmann2010extending} introduces distribution clauses so
-that some continuous distributions. Other work extended these methods
-further while focusing on efficient exact inference algorithms like
-model weight integration\cite{islam2012inference,
-  belle2015probabilistic}. Our work is most similar to
-\cite{gutmann2011magic} except that while they combine their forward
-reasoning with an importance sampler we use a particle cascade instead
-which can be more sample efficient.
+associating discrete distributions with facts. Early versions of
+ProbLog were also built on top of Prolog matching one of the goals of
+our work. Later work\cite{gutmann2010extending} introduces
+distribution clauses so that some continuous distributions. Other work
+extended these methods further while focusing on efficient exact
+inference algorithms like model weight
+integration\cite{islam2012inference, belle2015probabilistic}. Our work
+is most similar to \cite{gutmann2011magic} except that while they
+combine their forward reasoning with an importance sampler we use a
+particle cascade instead which can be more sample efficient.
 
@@
 
-\subsection{miniKanren}
+\subsection{microKanren}
 
-\textsc{miniKanren}\cite{byrd2017unified, daniel2018reasoned} is a
-pure logic programming language embedded in Scheme. The language
-consists of a set of terms, functions for constructing, a set of goal
-primitives, and two run functions to answer queries in the
-language. The goal primitives consist of a \texttt{fresh} form for
-introducing logic variables, a unification primitive, a conjunction
-combinator, and a disjunction combinator.
+\textsc{microKanren}\cite{10.1145/2989225.2989230, daniel2018reasoned}
+is a pure logic programming language embedded in Scheme. The language
+consists of a set of terms, a set of goal primitives, and two run
+functions to answer queries in the language. The goal primitives
+consist of a \texttt{fresh} form for introducing logic variables, a
+unification primitive, a conjunction combinator, a disjunction
+combinator, and ways to define and use relations.
 
-Define terms, goals, and substitutions here
+\subsection{Grammar and Definitions}
+
+\begin{figure}
+\begin{grammar}
+<prog> ::= (run <number> (<id>) <goal-expr>)
+
+<goal-expr> ::= (disj <goal-expr> <goal-expr>) \alt
+	        (conj <goal-expr> <goal-expr>) \alt
+                (fresh (<id>) <goal-expr>) \alt
+		(== <term-expr> <term-expr>) \alt
+		(letrec-rel ((<id> (<id> ...) <goal-expr>) ...) \\
+		\hspace{\grammarindent} <goal-expr>) \alt
+		(call-rel <lexical-var-ref> <term-expr> ...) \alt
+		(prim-rel-call <lexical-var-ref> <term-expr> ...) \alt
+		(delay <goal-expr>)
+
+<term-expr> ::= (quote <datum>) \alt
+                <lexical-var-ref> \alt
+                (cons <term-expr> <term-expr>)
+
+<term> ::= <number> \alt
+           \#f | \#t \alt
+	   <symbol> \alt
+	   (<term> . <term>) \alt
+	   <logic-var>
+
+\end{grammar}
+\end{figure}
+
+To create probkanren we extend this grammar with distribution clauses
+such as \texttt{normal} and \texttt{flip}. These are just another type
+of goal expression. We do not need to add a notion of probabilistic
+variables as they can just be seen as logic variables constrained in
+a particular way.
 
@@
 
-\section{Grammar and Definitions}
-
-Define distributional clauses here
-
 \section{Proposed Method}
 
-We propose to extend \textsc{miniKanren} by augmenting each of the
+We propose to extend \textsc{microKanren} by augmenting each of the
 search streams with a set of particles. These particles represent the
@@
 An initial set of particles is created from the probabilistic program
-when it is first run. As disjunctions like \texttt{conde} are encountered,
+when it is first run. As disjunctions (\texttt{disj}) are encountered,
 we split evenly the number of particles allocated to each stream. Whenever
@@
 Firstly, the arguments of distribution clauses must be
-ground. Secondly, a random variable can not unify with any arithmetic
-expression
+ground. Secondly, a random variable cannot unify with any arithmetic
+expression.
+
+This extension does not modify the search and streams are managed
+exactly as in microkanren. An additional advantage of this is thanks
+to the microkanren search being complete, if we generate enough
+particles we are guaranteed to recover the true posterior as all paths
+of the search space will eventually be explored.
 
@@
 
-One of the experiments should be something from the probabilistic logic literature.
+We validate that probKanren is at least as expressive as other probabilistic
+logic programming languages by implementing the Friends who Smoke model.
 
-Friends who Smoke is a probabilistic logic program which models the social nature of
-who smokes cigarettes. The model predicts that people who are friends with people
-who smoke are more likely to smoke.
+Friends who Smoke is a probabilistic logic program which models the
+social nature of who smokes cigarettes. The model predicts that people
+who are friends with people who smoke are more likely to smoke. We
+replicate the example on
+\url{https://dtai.cs.kuleuven.be/problog/tutorial/basic/05_smokers.html}
+using 2000 particles and get an empirical distribution that seems to
+match up with the discrete distribution returned from ProbLog.
 
-We could include experiment from Noah Goodman's dippl work like a semantic parsing
-example\\ \url{http://dippl.org/examples/semanticparsing.html}
+%% Compare to birch or Anglican on some top examples as well
 
-We could include program synthesis experiment, where probabilities
-allow us to specify a soft preference for using certain language
-primitives in a similar spirit to
-\textsc{rKanren}\cite{swords2013rkanren} or neural-guided search
-\cite{zhang2018neural}.
+%% We could include experiment from Noah Goodman's dippl work like a semantic parsing
+%% example\\ \url{http://dippl.org/examples/semanticparsing.html}
+
+%% We could include program synthesis experiment, where probabilities
+%% allow us to specify a soft preference for using certain language
+%% primitives in a similar spirit to
+%% \textsc{rKanren}\cite{swords2013rkanren} or neural-guided search
+%% \cite{zhang2018neural}.
 
@@
 
-We made a cool and simple to implement extension to
-\textsc{miniKanren} that let's us support probabilistic inference on
-both discrete and continuous distributions.
+We made a simple to implement extension to \textsc{microKanren} that
+let's us support probabilistic inference on both discrete and
+continuous distributions. The approach does not require modifying the
+underlying search algorithm or touch any of the backtracking code and
+comes with a theoretical guarantee that if the underlying search is
+complete then the probabilistic extension will require the true
+posterior given enough particles.
 

### 2021-09-14 paper/main.tex
Fix notation

@@
 \begin{abstract}
-  Probabilistic programming is sometimes described as a generalisation
+  Probabilistic programming can be conceptually seen as generalisation
   of logic programming where instead of just returning a set of
@@
   over those answers. But many contemporary probabilistic logic
-  programming languages are not implemented as simple extensions of
+  programming languages implementations are not simple extensions of
   existing logic programming languages but instead involve their own
@@
   programming language without needing to make any modifications to
-  the underlying search.
+  the underlying logic language's search. We use several illustrative
+  examples from the probabilistic programming and program synthesis
+  literature to demonstrate the practicality of the approach.
+
 \end{abstract}
@@
 
+%TODO: Examples should use the grammar defined in the paper
+
 \begin{minted}{Scheme}
@@
     (normal 0 3 q)
-    (greatero q 0))
+    (normal q 0 4))
 \end{minted}
@@
     (beta 1 1 q)
-    (flip q #t))
+    (bern q 1.0))
 \end{minted}
@@
 straight-line programs with no control-flow we can imagine numbering
-every sample function $x_1, f_2, \ldots, f_n$ and every observe
+every sample function $f_1, f_2, \ldots, f_n$ and every observe
 function $g_1, g_2, \ldots, g_n$ then our probability density over our
@@
   x_n &\sim& f(x_n \mid x_{0:n-1}) \\
-  w_n^{(k)} &=& \frac{g(x_n \mid x_{0:n})f(x_n \mid x_{0:n-1})}{q(x_n \mid x_{0:n-1})} \\
+  w_n^{(k)} &=& \frac{g(y_n \mid x_{0:n})f(x_n \mid x_{0:n-1})}{q(x_n \mid x_{0:n-1})} \\
   W_n^{(k)} &=& W_{n-1}^{(k)} w_n^{(k)}

### 2021-09-14 paper/main.tex
Use correct grammar in examples

@@
   (run 1000 (q)
+   (conj
     (normal 0 3 q)
-    (normal q 0 4))
+    (normal q 2 4))]
 \end{minted}
 
-\begin{minted}{Scheme}
-  (run 1000 (q)
-    (beta 1 1 q)
-    (bern q 1.0))
-\end{minted}
+%% \begin{minted}{Scheme}
+%%   (run 1000 (q)
+%%     (uniform 0 1 q)
+%%     (bern q 1.0))
+%% \end{minted}
 
@@
 consist of a \texttt{fresh} form for introducing logic variables, a
-unification primitive, a conjunction combinator, a disjunction
-combinator, and ways to define and use relations.
+unification primitive \texttt{==}, a conjunction combinator \texttt{conj}, a disjunction
+combinator \texttt{disj}, and ways to define and apply relations.
 
@@
                 <lexical-var-ref> \alt
-                (cons <term-expr> <term-expr>)
+                (cons <term-expr> <term-expr>) \alt
+                <term>
 
@@
 
-To create probkanren we extend this grammar with distribution clauses
-such as \texttt{normal} and \texttt{flip}. These are just another type
+To create \textsc{probKanren} we extend this grammar with distribution clauses
+such as \texttt{normal} and \texttt{bern}. These are just another type
 of goal expression. We do not need to add a notion of probabilistic
@@
 
+\section{Semantics of probKanren}
+
 \section{Experiments}

### 2021-09-14 paper/main.tex
Add semantics paragraph

@@
 \setminted{breaklines=true}
+
+\usepackage{cleveref}
 
@@
 consist of a \texttt{fresh} form for introducing logic variables, a
-unification primitive \texttt{==}, a conjunction combinator \texttt{conj}, a disjunction
-combinator \texttt{disj}, and ways to define and apply relations.
+unification primitive \texttt{==}, a conjunction combinator
+\texttt{conj}, a disjunction combinator \texttt{disj}, and ways to
+define and apply relations. Further forms are shown in detail in
+\Cref{fig:grammar}.
 
@@
 \begin{figure}
+  \centering
   \begin{grammar}
@@
   \end{grammar}
+  \caption{Grammar for \textsc{microKanren}}
+  \label{fig:grammar}
 \end{figure}
@@
 
-Sequential Monte Carlo\cite{chopin2020introduction} are an efficient
+Sequential Monte Carlo\cite{chopin2020introduction}(SMC) is an efficient
 online way to sample from probabilistic models especially suited for
@@
 
-We propose to extend \textsc{microKanren} by augmenting each of the
-search streams with a set of particles. These particles represent the
+We propose to extend \textsc{microKanren} by turning the search into a
+SMC sampler. We accomplish this by augmenting each of the search
+streams with a set of particles. These particles represent the
 empirical distribution of that stream. Each particle has associated
 with it a substitution of all the logic and random variables as well
-as a weight that is proportional to the likelihood of the substitution.
+as a weight that is proportional to the likelihood of the
+substitution.
 
@@
 \section{Semantics of probKanren}
+
+We follow the semantics of \cite{staton2016semantics}, where each
+language form has a denotational semantics that associates a
+measurable function with each form, and an operational semantics that associates
+a sampler with each form.
 

### 2021-09-14 paper/main.tex
Fix description of program
Add more on microkanren

@@
 
-This probKanren program draws 1000 samples from a normal distribution truncated
-below 0.
+This probKanren program draws 1000 samples from a conditional normal distribution.
 
@@
 \Cref{fig:grammar}.
+
+Goal expressions represent a possibly backtracking search
+procedure. These goals all take as input some state and output a
+stream of possible output states. We run these goals by starting with
+an initial state that holds an empty substitution dictionary, and
+passing it into the top-level goal expression which then passes it
+recursively down to sub-expressions. Encountering \texttt{fresh} adds
+a new logical variable to the substitution; \texttt{conj} will apply
+the first goal to the state passed in, and then apply the second goal
+to the states associated with each resulting steam finally unioning
+the results; \texttt{disj} will apply both goals to the same passed in
+state and concatenate the resulting steam; and \texttt{==} unifies its
+arguments in the context of the current state discarding any streams
+which fail to unify.
+
+To handle goal expressions that might diverge, the streams are
+expanded in an interleaving\cite{kiselyov2005backtracking} fashion
+giving each branch a chance to produce answers. This interleaving search
+is sound and complete\cite{rozplokhas2019certified}.
 
@@
 If we imagine running this program each execution trace can be seen as
-a sample from the distribution. If we then weight by their likelihood
-as we run them that collection execution traces will be an empirical
+a sample from the distribution. If we then weigh by their likelihood
+as we run them that collection of execution traces will be an empirical
 distribution of the given program. We call each of these execution traces

### 2021-09-15 paper/main.tex
Fix SMC bits

@@
 
-This probKanren program draws 1000 samples from a conditional normal distribution.
+In the above program, we have a probabilistic model for the Gaussian
+unknown mean problem. We observe a value 4 from the $\mathcal{N}(q,2)$
+distribution, and that $q$ has a prior distribution of $\mathcal{N}(0, 3)$.
+This \textsc{probKanren} program draws 1000 samples from a conditional normal distribution
+that represents the posterior probability distribution associated with $q$.
 
@@
 
-To create \textsc{probKanren} we extend this grammar with distribution clauses
-such as \texttt{normal} and \texttt{bern}. These are just another type
-of goal expression. We do not need to add a notion of probabilistic
-variables as they can just be seen as logic variables constrained in
-a particular way.
+To create \textsc{probKanren} we extend this grammar with distribution
+clauses such as \texttt{normal} and \texttt{bern}. Distribution clauses
+take as arguments the parameters of the distribution and a last argument
+representing a draw from that distribution. For example, in \texttt{(normal 0 1 x)}
+means \texttt{x} represents a draw from $\mathcal{N}(0,1)$ distribution.
+
+These are just another type of goal expression. We do not need to add
+a notion of probabilistic variables to the language as they can
+treated as logic variables constrained in a particular way. The
+semantics of the language though does change from a set of answers to
+a query to a probability distribution of answers to a query.
+
+We follow the semantics of \cite{staton2016semantics}, where each
+language form has a denotational semantics that associates a
+measurable function with each form, and an operational semantics that associates
+a sampler with each form.
 
@@
 function $g_1, g_2, \ldots, g_n$ then our probability density over our
-random variables $x$ and observed data $y$ can be defined as:
+random variables $x_{0:T}$ and observed data $y_{0:T}$ can be defined as:
 
 \begin{equation}
-  p(x, y) = \prod_i f(x_i \mid x_{0:n-1})g(y_i \mid x_{0:n})
+  p(x_{0:T}, y_{0:T}) = p(x_0)\prod_{t=1}^T f_t(x_t \mid x_{t-1})\prod_{t=0}^T g_t(y_t \mid x_t)
 \end{equation}
@@
 
-If we imagine running this program each execution trace can be seen as
-a sample from the distribution. If we then weigh by their likelihood
-as we run them that collection of execution traces will be an empirical
-distribution of the given program. We call each of these execution traces
-particles and the below method is how we obtain them along with their weights $w$.
+The simplest way to sample from our target distribution $p(x_{0:T}
+\mid y_{0:T})$ is to samples from a sequence of intermediary
+distributions $p(x_{0:t} \mid y_{0:t})$ where $t$ goes from $1$ to
+$T$. We do this by sampling a population of $N$ particles
+$\{x^{(1)}_t, \ldots, x^{(i)}_t, \ldots x^{(N)}_t\}$ from a proposal
+distribution $q(x_t \mid x_{0:t-1}, y_{0:t})$, which when combined
+with a set of importance weights $\{w^{(1)}_t, \ldots, w^{(i)}_t,
+\ldots w^{(N)}_t\}$ lets us approximate each intermediary
+distribution.
 
-\begin{eqnarray}
-  x_n &\sim& f(x_n \mid x_{0:n-1}) \\
-  w_n^{(k)} &=& \frac{g(y_n \mid x_{0:n})f(x_n \mid x_{0:n-1})}{q(x_n \mid x_{0:n-1})} \\
-  W_n^{(k)} &=& W_{n-1}^{(k)} w_n^{(k)}
-\end{eqnarray}
+Thanks to the recurrence relation
 
-The weights are then normalised at the end to make a proper probability distribution.
+\begin{equation}
+p(x_{0:t} \mid y_{0:t}) = p(x_{0:t-1} \mid y_{0:t-1}) \frac{f_t(x_t \mid x_{t-1}) g_t(y_t \mid x_t)}{p(y_t \mid y_{0:t-1})}
+\end{equation}
+
+We know in $T$ rounds we compute the desired distribution as follows:
+
+Initialise with:
+
+\begin{align*}
+x_0^{(i)} &\sim p(x_0) \\
+w_0^{(i)} &= 1/N
+\end{align*}
+
+Then for $t$ from $1$ to $T$
+
+\begin{align*}
+  x_t^{(i)} &\sim q(x_t \mid x^{(i)}_{0:t-1}) \\
+  w_t^{(i)} &\propto w_{t-1}^{(i)} \frac{g(y_t \mid x^{(i)}_t)f(x^{(i)}_t \mid x^{(i)}_{t-1})}{q(x^{(i)}_t \mid x^{(i)}_{0:t-1})}
+\end{align*}
+
+We get the best results if $q(x_t \mid x_{0:t-1})$ is equal to $p(x_t \mid x_{0:t-1}, y_{0:t})$.
 
 \subsection{Sequential Importance Resampling}
+
+Unfortunately, SIS is likely to over time lead to most of the weight
+in a small fraction of the particles with the rest of the particles
+having negligible weight. This is sometimes called \emph{weight degeneracy}. We
+address this by replicating the particles with high weight and
+dropping the ones with low weight. This is called a resampling step,
+and it happens after each round.
+
+\begin{align*}
+  a_t^{(i)} &\sim r(w_{t-1}^{(1:N)}) \\
+  x_t^{(i)} &\sim q(x_t \mid x^{(a_t^{(i)})}_{0:t-1}) \\
+  w_t^{(i)} &\propto w_{t-1}^{(i)} \frac{g(y_t \mid x^{(i)}_t)f(x^{(i)}_t \mid x^{(i)}_{t-1})}{q(x^{(i)}_t \mid x^{(i)}_{0:t-1})}
+\end{align*}
 
@@
 
-As the SMC resampling step was defined in the previous section
-Particle Cascades \cite{PaigeWDT14} remove this barrier allowing
-every particle to be resampled asynchronously with the associated
-weights being relative to a global running average. % double-check this
+Unfortunately, SMC as defined here requires having access to all the
+particles at every step in the process. This conflicts with the
+functional nature of our implementation. Instead we make use of
+Particle Cascades \cite{PaigeWDT14}, remove this barrier allowing every
+particle to be resampled asynchronously with the associated weights
+being relative to a global running average. % double-check this
 
@@
 
-An initial set of particles is created from the probabilistic program
-when it is first run. As disjunctions (\texttt{disj}) are encountered,
-we split evenly the number of particles allocated to each stream. Whenever
-we encounter a unification primitive, we run a resampling step. This
-helps to prune low-weight particles and replicate high weight ones.
-
-As an optimisation we may create more particles during resampling
-based on a globally stored a counter of the effective sample size of
-all particles across all streams.
-
 We follow \cite{gutmann2010extending} and place the following restrictions
@@
 
+An initial set of particles is created from the probabilistic program
+when it is first run. When primitives such as \texttt{normal} we first
+check is all the parameter terms are grounded and then if the last
+argument is fresh, we sample a value for it for each particle and then
+add this sampled value along with the associated logic variable to the
+substitution associated with that particle. If all the terms are ground,
+we treat the primitive like an observation statement and multiply the
+weights of each substitution with the likelihood of the observation.
+
+As disjunctions (\texttt{disj}) are encountered, we split evenly the
+number of particles allocated to each stream. Whenever we encounter a
+unification primitive, we run a resampling step. This helps to prune
+low-weight particles and replicate high weight ones.
+
+As an optimisation we may create more particles during resampling
+based on a globally stored a counter of the effective sample size of
+all particles across all streams.
+
 This extension does not modify the search and streams are managed
@@
 of the search space will eventually be explored.
-
-\section{Semantics of probKanren}
-
-We follow the semantics of \cite{staton2016semantics}, where each
-language form has a denotational semantics that associates a
-measurable function with each form, and an operational semantics that associates
-a sampler with each form.
 

### 2021-09-15 paper/main.tex
Give a definition for r

@@
 
-The above is called multinomial resampling but there are other methods
-as well. A survey\cite{douc2005comparison} of resampling methods
-suggests all of them are helpful to reduce particle degeneracy.
+We set $r(w) = \text{Categorical}(w)$, which is called multinomial resampling but there are other methods
+as well\cite{douc2005comparison}.
 

### 2021-09-15 paper/main.tex
Fix labels
More fixes

@@
 %% This command is for the conference information
-\conference{PLP '21: The 8th Workshop on Probabilistic Logic Programming,
-  September 20--27, 2021, Porto, Portugal}
+\conference{PLP 2021}
 
@@
   existing logic programming languages but instead involve their own
-  unique implementations. Here we introduce probKanren, a simple
+  unique implementations. Here we introduce \textsc{probKanren}, a simple
   extension to microKanren that transforms it into a probabilistic
@@
 
-But the conceptual simplicity of this sort of generalisation is not reflected
-in the complexity of many existing probabilistic logic programming systems.
-They often involve making implementing sophisticated algorithms and the underlying
-systems are not just implemented on top of existing logic programming systems.
+But the conceptual simplicity of this sort of generalisation is not
+reflected in the complexity of many existing probabilistic logic
+programming systems.  They often involve making implementing
+sophisticated inference algorithms and the underlying systems are not
+just implemented on top of existing logic programming systems.
 
-We believe a conceptually simple extension deserves a conceptually
-simple implementation to go along with it. We thus contribute a simple
-way to extend microkanren a small logic programming DSL such that it
-becomes a probabilistic programming language.
+We believe that a conceptually simple extension deserves a
+conceptually simple implementation to go along with it. We thus
+contribute a simple way to extend microkanren a small logic
+programming domain-specific language such that it becomes a
+probabilistic programming language.
 
@@
 
-To help explain how to use probKanren we introduce the following example:
+To help explain how to use \textsc{probKanren} we introduce the following example:
 
@@
     (normal 0 3 q)
-    (normal q 2 4))]
+    (normal q 2 4)))
 \end{minted}
@@
 associating discrete distributions with facts. Early versions of
-ProbLog were also built on top of Prolog matching one of the goals of
+ProbLog were also built on top of Prolog, matching one of the goals of
 our work. Later work\cite{gutmann2010extending} introduces
-distribution clauses so that some continuous distributions. Other work
-extended these methods further while focusing on efficient exact
-inference algorithms like model weight
-integration\cite{islam2012inference, belle2015probabilistic}. Our work
-is most similar to \cite{gutmann2011magic} except that while they
-combine their forward reasoning with an importance sampler we use a
-particle cascade instead which can be more sample efficient.
+distribution clauses so that some continuous distributions like
+Gaussians can be represented. Other work extended these methods
+further while focusing on efficient exact inference algorithms like
+weighted model integration\cite{islam2012inference,
+  belle2015probabilistic}. Our work is most similar to
+\cite{gutmann2011magic}, except that while they combine their forward
+reasoning with an importance sampler we use a particle cascade instead
+which can be more sample efficient.
 
@@
 consists of a set of terms, a set of goal primitives, and two run
-functions to answer queries in the language. The goal primitives
+functions that execute queries in the language. The goal primitives
 consist of a \texttt{fresh} form for introducing logic variables, a
@@
 the first goal to the state passed in, and then apply the second goal
-to the states associated with each resulting steam finally unioning
-the results; \texttt{disj} will apply both goals to the same passed in
-state and concatenate the resulting steam; and \texttt{==} unifies its
-arguments in the context of the current state discarding any streams
-which fail to unify.
+to the states associated with each resulting stream from the first
+goal and finally unify the results; \texttt{disj} will apply both
+goals to the same input state and concatenate the resulting streams;
+and \texttt{==} unifies its arguments in the context of the current
+state, discarding any streams which fail to unify.
 
 To handle goal expressions that might diverge, the streams are
-expanded in an interleaving\cite{kiselyov2005backtracking} fashion
+expanded in an interleaving\cite{kiselyov2005backtracking} fashion,
 giving each branch a chance to produce answers. This interleaving search
@@
 
-To create \textsc{probKanren} we extend this grammar with distribution
+To create \textsc{probKanren}, we extend this grammar with distribution
 clauses such as \texttt{normal} and \texttt{bern}. Distribution clauses
@@
 representing a draw from that distribution. For example, in \texttt{(normal 0 1 x)}
-means \texttt{x} represents a draw from $\mathcal{N}(0,1)$ distribution.
+means \texttt{x} represents a draw from the $\mathcal{N}(0,1)$ distribution.
 
 These are just another type of goal expression. We do not need to add
-a notion of probabilistic variables to the language as they can
-treated as logic variables constrained in a particular way. The
-semantics of the language though does change from a set of answers to
-a query to a probability distribution of answers to a query.
+a notion of probabilistic variables to the language grammar, as they
+can treated as logic variables constrained in a particular way. The
+semantics of the language though does change from an answer set to a
+probability distribution on that answer set.
 
 We follow the semantics of \cite{staton2016semantics}, where each
-language form has a denotational semantics that associates a
-measurable function with each form, and an operational semantics that associates
-a sampler with each form.
+language form has the denotational semantics that a
+measurable function, and an operational semantics of a sampler.
 
@@
 Probabilistic Programming Languages\cite{wood2014new,
-  van2018introduction} are a family of domain specific languages for
+  van2018introduction} are a family of domain-specific languages for
 posing and efficiently solving probabilistic modelling problems. At
-their core, all have a way to \texttt{sample} from a probability
-distribution and \texttt{observe} data generated from a probability
-distribution.
+their core, all have a way to \texttt{sample} and \texttt{observe}
+data under a probability distribution or Markov kernel.
 
-There are many ways to implement inference algorithms for probabilistic
-programming languages but methods based on likelihood-weighting and
-sequential monte carlo algorithms are the easiest.
+There are many ways to inference algorithms for probabilistic
+programming languages, but methods based on likelihood-weighting and
+Sequential Monte Carlo algorithms are the simplest to implement.
 
@@
 Sequential Monte Carlo\cite{chopin2020introduction}(SMC) is an efficient
-online way to sample from probabilistic models especially suited for
+online way to sample from probabilistic models, which is especially suited for
 state-space domains. If we imagine our probabilistic programs as
@@
 function $g_1, g_2, \ldots, g_n$ then our probability density over our
-random variables $x_{0:T}$ and observed data $y_{0:T}$ can be defined as:
+latent variables $x_{0:T}$ and observed data $y_{0:T}$ can be defined as:
 
@@
 The simplest way to sample from our target distribution $p(x_{0:T}
-\mid y_{0:T})$ is to samples from a sequence of intermediary
+\mid y_{0:T})$ is to sample from a sequence of intermediary
 distributions $p(x_{0:t} \mid y_{0:t})$ where $t$ goes from $1$ to
@@
 
-Thanks to the recurrence relation
+Thanks to the recurrence relation:
 
@@
 
-\begin{align*}
+\begin{align}
 x_0^{(i)} &\sim p(x_0) \\
 w_0^{(i)} &= 1/N
-\end{align*}
+\end{align}
 
@@
 
-\begin{align*}
+\begin{align}
   x_t^{(i)} &\sim q(x_t \mid x^{(i)}_{0:t-1}) \\
-  w_t^{(i)} &\propto w_{t-1}^{(i)} \frac{g(y_t \mid x^{(i)}_t)f(x^{(i)}_t \mid x^{(i)}_{t-1})}{q(x^{(i)}_t \mid x^{(i)}_{0:t-1})}
-\end{align*}
+  w_t^{(i)} &\propto w_{t-1}^{(i)} \frac{f(x^{(i)}_t \mid x^{(i)}_{t-1})g(y_t \mid x^{(i)}_t)}{q(x^{(i)}_t \mid x^{(i)}_{0:t-1})}
+\end{align}
 
@@
 
-Unfortunately, SIS is likely to over time lead to most of the weight
-in a small fraction of the particles with the rest of the particles
-having negligible weight. This is sometimes called \emph{weight degeneracy}. We
-address this by replicating the particles with high weight and
+Unfortunately, over time SIS is likely to lead to most of the
+importance weight in a small fraction of the particles, with the rest
+of the particles having negligible weight. This is usually called
+\emph{weight degeneracy}. We address this sample efficiency issue in
+the standard way by replicating the particles with high weight and
 dropping the ones with low weight. This is called a resampling step,
-and it happens after each round.
+and it happens after each round. Resampling effectively removes
+particles with low weight and duplicates particles with higher weight
+by sampling with replacement our existing particles.
 
@@
   x_t^{(i)} &\sim q(x_t \mid x^{(a_t^{(i)})}_{0:t-1}) \\
-  w_t^{(i)} &\propto w_{t-1}^{(i)} \frac{g(y_t \mid x^{(i)}_t)f(x^{(i)}_t \mid x^{(i)}_{t-1})}{q(x^{(i)}_t \mid x^{(i)}_{0:t-1})}
+  w_t^{(i)} &\propto w_{t-1}^{(i)} \frac{f(x^{(i)}_t \mid x^{(i)}_{t-1})g(y_t \mid x^{(i)}_t)}{q(x^{(i)}_t \mid x^{(i)}_{0:t-1})}
 \end{align*}
 
-The problem with the above algorithm is over time for many particles
-$W_k$ is going to become low and that particle will stop being very
-informative of the underlying distribution. To mitigate this issue,
-each time we encounter an observation we resample our particles. Resampling
-effectively removes particles with low weight and duplicates particles
-with higher weight by sampling with replacement our existing particles.
-
-We set $r(w) = \text{Categorical}(w)$, which is called multinomial resampling but there are other methods
-as well\cite{douc2005comparison}.
+For each particle $i$, resampling tells us which of the existing
+particles $a_i$ will be its ancestor. We set $r(w) =
+\text{Categorical}(w)$, which is called multinomial resampling but
+there are other methods as well\cite{douc2005comparison}.
 
@@
 functional nature of our implementation. Instead we make use of
-Particle Cascades \cite{PaigeWDT14}, remove this barrier allowing every
+Particle Cascades \cite{PaigeWDT14}, removing this barrier allowing every
 particle to be resampled asynchronously with the associated weights
@@
 
-We propose to extend \textsc{microKanren} by turning the search into a
+We propose to extend \textsc{microKanren} by turning the search into an
 SMC sampler. We accomplish this by augmenting each of the search
@@
 Firstly, the arguments of distribution clauses must be
-ground. Secondly, a random variable cannot unify with any arithmetic
+grounded. Secondly, a random variable cannot unify with any arithmetic
 expression.
@@
 An initial set of particles is created from the probabilistic program
-when it is first run. When primitives such as \texttt{normal} we first
-check is all the parameter terms are grounded and then if the last
-argument is fresh, we sample a value for it for each particle and then
-add this sampled value along with the associated logic variable to the
-substitution associated with that particle. If all the terms are ground,
-we treat the primitive like an observation statement and multiply the
-weights of each substitution with the likelihood of the observation.
+when it is first run. When encountering primitives such as
+\texttt{normal} we first check if all the parameter terms are grounded
+and then if the last argument is fresh, we sample a value for it for
+each particle and then add this sampled value along with the
+associated logic variable to the substitution associated with that
+particle. If all the terms are ground, we treat the primitive like an
+observation statement and multiply the weights of each substitution
+with the likelihood of the observation.
+
+As conjunctions (\texttt{conj}) are encountered, all particles in all
+output streams from the first goal become part of the initial states
+that are each passed to the second goal.
 
@@
 unification primitive, we run a resampling step. This helps to prune
-low-weight particles and replicate high weight ones.
+particles with low weight and replicate particles with high weight.
 
-As an optimisation we may create more particles during resampling
-based on a globally stored a counter of the effective sample size of
+As an optimisation, we may create more particles during resampling
+based on a globally stored counter of the effective sample size of
 all particles across all streams.
@@
 This extension does not modify the search and streams are managed
-exactly as in microkanren. An additional advantage of this is thanks
-to the microkanren search being complete, if we generate enough
-particles we are guaranteed to recover the true posterior as all paths
-of the search space will eventually be explored.
+exactly as in \textsc{microKanren}. An additional advantage, due to
+the completeness of the \text{microKanren} search being complete, is
+that we are guaranteed to recover the true posterior as all paths of
+the search space will eventually be explored given enough particles.
 
@@
 
-We validate that probKanren is at least as expressive as other probabilistic
+We validate that \textsc{probKanren} is at least as expressive as other probabilistic
 logic programming languages by implementing the Friends who Smoke model.
@@
 We made a simple to implement extension to \textsc{microKanren} that
-let's us support probabilistic inference on both discrete and
+lets us support probabilistic inference on both discrete and
 continuous distributions. The approach does not require modifying the
-underlying search algorithm or touch any of the backtracking code and
+underlying search algorithm or touching any of the backtracking code and
 comes with a theoretical guarantee that if the underlying search is
-complete then the probabilistic extension will require the true
+complete then the probabilistic extension will recover the true
 posterior given enough particles.

### 2021-09-18 paper/main.tex
Misc fixes

@@
 reflected in the complexity of many existing probabilistic logic
-programming systems.  They often involve making implementing
-sophisticated inference algorithms and the underlying systems are not
-just implemented on top of existing logic programming systems.
+programming systems.  They often involve implementing sophisticated
+inference algorithms and the underlying systems are not just
+implemented on top of existing logic programming systems.
 
@@
 conceptually simple implementation to go along with it. We thus
-contribute a simple way to extend microkanren a small logic
+contribute a simple way to extend \textsc{microKanren}, a small logic
 programming domain-specific language such that it becomes a
@@
 consists of a set of terms, a set of goal primitives, and two run
-functions that execute queries in the language. The goal primitives
+forms that turn goal queries into a set of answers. The goal primitives
 consist of a \texttt{fresh} form for introducing logic variables, a
@@
 passing it into the top-level goal expression which then passes it
-recursively down to sub-expressions. Encountering \texttt{fresh} adds
-a new logical variable to the substitution; \texttt{conj} will apply
-the first goal to the state passed in, and then apply the second goal
-to the states associated with each resulting stream from the first
-goal and finally unify the results; \texttt{disj} will apply both
-goals to the same input state and concatenate the resulting streams;
-and \texttt{==} unifies its arguments in the context of the current
-state, discarding any streams which fail to unify.
+recursively down to sub-expressions. Encountering \texttt{fresh}
+extends the lexical environment with a binding between a lexical
+variable and a logic variable; \texttt{conj} will apply the first goal
+to the state passed in, and then apply the second goal to the states
+associated with each resulting stream from the first goal and finally
+unify the results; \texttt{disj} will apply both goals to the same
+input state and concatenate the resulting streams; and \texttt{==}
+unifies its arguments in the context of the current state, discarding
+any streams which fail to unify.
 
@@
 take as arguments the parameters of the distribution and a last argument
-representing a draw from that distribution. For example, in \texttt{(normal 0 1 x)}
+representing a draw from that distribution. For example, \texttt{(normal 0 1 x)}
 means \texttt{x} represents a draw from the $\mathcal{N}(0,1)$ distribution.
@@
 
-We follow the semantics of \cite{staton2016semantics}, where each
-language form has the denotational semantics that a
+We follow the semantics of \cite{staton2016semantics}, this is a
+denotational semantics where each language form is associated with a
 measurable function, and an operational semantics of a sampler.
@@
 
-There are many ways to inference algorithms for probabilistic
-programming languages, but methods based on likelihood-weighting and
-Sequential Monte Carlo algorithms are the simplest to implement.
+There are many inference algorithms for probabilistic programming
+languages, but methods based on likelihood-weighting and Sequential
+Monte Carlo algorithms are the simplest to implement.
 
@@
 
-We know in $T$ rounds we compute the desired distribution as follows:
+we know in $T$ rounds we compute the desired distribution as follows:
 
@@
 exactly as in \textsc{microKanren}. An additional advantage, due to
-the completeness of the \text{microKanren} search being complete, is
-that we are guaranteed to recover the true posterior as all paths of
-the search space will eventually be explored given enough particles.
+the \text{microKanren} search being complete, we are guaranteed to
+recover the true posterior as all paths of the search space will
+eventually be explored given enough particles.
 
@@
 
-We made a simple to implement extension to \textsc{microKanren} that
+We made a simple-to-implement extension to \textsc{microKanren} that
 lets us support probabilistic inference on both discrete and
